### 2017-05-06 paper.tex
initial word/brain dump

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{%
The Nuclear Forensics Problem and Statistical Methods: \\
Evaluating the Nuclear Fuel Cycle with Machine Learning
}
\author{Arrielle C. Opotowsky, Paul P.H. Wilson}

\institute{
Computational Nuclear Engineering Research Group \\
University of Wisconsin at Madison
}

\email{opotowsky@wisc.edu \and paulwilson@wisc.edu}

% Optional disclaimer: remove this command to hide
% \disclaimer{Notice: }

%%%% packages and definitions (optional)
\usepackage{graphicx} % allows inclusion of graphics
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\grad}{\vec{\nabla}} % gradient
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

In the event of a nuclear incident, such as the retrieval of stolen nuclear
material or the detonation of a dirty bomb, it is necessary to learn as much as
possible about the source of the materials in a timely manner. To characterize
the materials, both radiological methods (e.g., gamma spectroscopy) and
ionization methods (e.g., mass spectroscopy) are used to determine isotopic
ratios, chemical compounds, or trace elements. Although each category has a
multitude of techniques within it, the main tradeoff is between time/cost and
amount of information gained.

The results of these analytic techniques are then compared against existing
databases to determine the origin of the nuclear material(s). These databases
are highly multidimensional, and furthermore, are rife with missing data
entries and inconsistent uncertainties. Direct comparison between measurement
results and a database therefore may not yield accurate results. Fortunately,
machine learning algorithms can be explored to create a model from a database
to "fill between the lines".  Additionally, having a machine-learned model may
overcome the above challenges of multidimensionality, missing data, and
irregular uncertainty.

While different machine learning algorithms and parameters will be
investigated, it is first important to determine if statistical methods can
overcome the inherent database deficiencies. Thus, this paper focuses on
probing the amount of information required to obtain realistic results.  This
can be best understood as the analgous real-world scenario.  Although mass
spectroscopy techniques provide extremely accurate isotopic information, they
are time-consuming and more expensice. And although gamma spectroscopy can give
extremely fast results cheaply, it only measures certain radiological signals
and is influenced by many environmental (or storage) factors. In the simulation
and machine learning paradigm, we need to determine what exactly is needed to
train a machine-learned model. Can the algorithm overcome the deficiencies of
gamma detection and still provide useful results? Or does it need more
information, e.g., exact isotopics?

Thus, ultimately, the goal is to answer the question \textit{How does the
ability to determine forensic-relevant spent nuclear fuel attributes degrade as
less information is available?}. But first, we must establish some baseline
expectations and algorithms to use. This work is based off previous work on the
subject (cite Dayman), and expands upon it by also evaluating a more advanced
machine learning algorithm: neural nets. Below is a more in depth discussion of
nuclear forensics and how machine learning can contribute to this research
area. After that, an experimental design is outlined. Lastly, the results are
presented and discussed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background and Theory}

\subsection{Nuclear Forensics}

Nuclear forensics is the analysis and interpretation to determine the history
of nuclear material, whether that be intercepted spent nuclear fuel, uranium
ore concentrate, or the debris from an exploded nuclear device. Technical
nuclear forensics focuses on the characterization and interpretation of those
results. This, in combination with intelligence data, aids in material
attribution, which is the overall goal of nuclear forensics.

(Something on goal to get as much info as possible within 24 h of interception
or device detonation).

Nuclear Forensics Workflow Although this focuses on Some materials of special
interest: SNF and reprocessed MOX.  For SNF, measure material, use isotope
content and/or isotope ratios to determine things like reactor type, fuel type
and enrichment at beginning of irradiation, cooling time, burnup. For
reprocessed MOX, ??.

After the material characteristics are measured, they are matched in a
forensics database that includes some or all this information for
pre-existing/pre-measured SNF/MOX. These databases are kept by individual
countries, and a given database will have widely varying uncertainty depending
where the material was measured as well as missing data in some fields.
Therefore, matching can be difficult.

(Classification, Characterization, Interpretation (Analysis), Reconstruction
(Attribution) - from the New Nuclear Forensics book) (Char methods to get
isotopic ratios or use S/ML, Interp examples)

To accomplish this, (we want to see if) it is possible to limit material
measurements to rapid-result nondestructive analysis, such as gamma ray
spectroscopy. Gamma spec gives less information at a higher uncertainty than
the near perfect results of some destructive mass spec techniques, like TIMS.
Additionally, within gamma spec techniques (field vs lab), uncertainties can
also vary.

\subsection{Machine Learning}

Given imperfect data with varying amounts of uncertainty as well as the
required comparison to imperfect databases, many have begun considering
artificial intelligence approaches to nuclear forensics problems, such as
implementing searching algorithms for database comparison and machine learning
for determining spent fuel characteristics (cite all). A variety of statistical
and machine learning tools have been used to both classify spent fuel (reactor
type, fuel type) and predict values such as burnup, initial enrichment, or
cooling time (regression).

Add some real ML background here
\begin{subequations} \label{eqs:fullTransport}
\begin{multline} \label{eq:fullTransportVol}
  \vec{\Omega}\vd \grad \psi(\vec{x}, \vec{\Omega})
  + \sigma(\vec{x}) \psi (\vec{x}, \vec{\Omega})
\\ =
  \frac{\sigma_s(\vec{x})}{4\pi} \int_{4\pi} \psi(\vec{x},\vec{\Omega}')
  \ud\Omega' + \frac{q(\vec{x})}{4\pi}
  \equiv \frac{1}{4\pi} Q(\vec{x}) \,,
\end{multline}
inside $\vec{x} \in V$, $\vec{\Omega} \in 4\pi$, with an incident boundary
condition
\begin{equation} \label{eq:fullTransportBndy}
  \psi(\vec{x}, \vec{\Omega}) = \psi^b(\vec{x}, \vec{\Omega}) \,,
 \quad \vec{x} \in \partial V, \ \vec{\Omega} \vd \vec{n} < 0\,.
\end{equation}
\end{subequations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Design}

Talk about Dayman paper and how I'm extending that, to include neural nets
first and to eval mucking up the training data second.

\subsection{Algorithms Used}

Preprocessing, kNN, ridge, neural net - these are all discussed above, but can
talk deets + parameters here

\subsection{Validation of Each Algorithm}

Classification Training Error: Predetermined test set for training error,
perhaps can make a learning curve (maybe not until there is a variable
train/test set for cross validation)

Classification and Regression Error Confidence: Confidence intervals on error
(helps understand true error versus just the sample error), Test set must be >
30 instances, Can easily calculate N\% confidence interval.

Generalizability should be evaluated (learning more about this with cs760
project)

\subsection{Comparing Algorithms}

Options for comparison of algorithms: Comparing classification of 2 classes on
same ROC plot with multiple ML systems, Scatter plots, Pairwise t-tests

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
The results were interesting, so interesting in fact that we have decided to
present them here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection Goes Here}
The user must manually capitalize initial letters of a subsection heading.

For those who like equations in their papers, \LaTeX\ is a good choice. Here is
an equation for the Marshak diffusion boundary condition:
\begin{equation} \label{eq:marshak}
  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
\end{equation}
If we so choose, we can effortlessly reference the equation later.

Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
vacuum boundary condition:
\begin{equation*}
  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
\end{equation*}
The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
an extrapolation distance of about $0.71045$.

Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
document. Always place figures after they are referenced so as not to throw
off the reader. You can use symbols and different line styles to help
differentiate your results, especially if they are printed in black and white.
Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
solution, solid lines \verb|-| for the new method's solutions, and dotted lines
\verb|:| for existing inaccurate methods.
\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics{example_figure}
  \caption{Captions are flush with the left.}
  \label{fig:voltage}
\end{figure}

Later on, we can include a table, even one that spans two columns such as
Table~\ref{tab:widetable}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htb]
  \centering
\begin{tabular}{llllllllll}\toprule
      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
      $\varepsilon$      & $N_\text{it}$
\\ \midrule
$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
\\
$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
\\
$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
\\
$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
\\
\bottomrule
\end{tabular}
  \caption{This is an example of a really wide table which might not normally
  fit in the document.}
  \label{tab:widetable}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Notice how the table reference uses a Roman numeral
for its numbering scheme, whereas the figure reference uses an Arabic numeral.
For one-column tables, use the \verb|table| environment; two-column tables use
\verb|table*|. The same applies to figures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Another Subsection}
Excessive sectioning in a three-page document is discouraged, but here are more
subsections to demonstrate compliance with the ANS formatting guidelines.

\subsubsection{Third-level Heading}
This subsubsection shows compliance with the ANS-specified standard. This level
of heading should be used rarely.

\subsubsection{Another Such Heading}
And, if you really think you need a third-level heading, you should make sure
that your subsection needs at least two of them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The included ANS style file and this clear example file are a panacea for
the hours of headache that invariably results from formatting a document in
Microsoft Word.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Appendix}

Numbering in the appendix is different:
\begin{equation} \label{eq:appendix}
  2 + 2 = 5\,.
\end{equation}
and another equation:
\begin{equation} \label{eq:appendix2}
  a + b = c\,.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This material is based upon work supported a Department of Energy Nuclear
Energy University Programs Graduate Fellowship.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{bibliography}
\end{document}

### 2017-05-18 paper.tex
adding some experimental design notes

@@
 The Nuclear Forensics Problem and Statistical Methods: \\
-Evaluating the Nuclear Fuel Cycle with Machine Learning
+Evaluating Machine Learning for Spent Fuel Prediction
 }
@@
 
-(Something on goal to get as much info as possible within 24 h of interception
-or device detonation).
-
-Nuclear Forensics Workflow Although this focuses on Some materials of special
-interest: SNF and reprocessed MOX.  For SNF, measure material, use isotope
-content and/or isotope ratios to determine things like reactor type, fuel type
-and enrichment at beginning of irradiation, cooling time, burnup. For
-reprocessed MOX, ??.
+Workflow for determining SNF quantities of interest: measure material, use
+isotope content and/or isotope ratios to determine things like reactor type,
+fuel type and enrichment at beginning of irradiation, cooling time, burnup.
 
@@
 forensics database that includes some or all this information for
-pre-existing/pre-measured SNF/MOX. These databases are kept by individual
+pre-existing/pre-measured SNF. These databases are kept by individual
 countries, and a given database will have widely varying uncertainty depending
@@
 
+(Something on goal to get as much info as possible within 24 h of interception
+or device detonation).
+
 \subsection{Machine Learning}
@@
 Add some real ML background here
-\begin{subequations} \label{eqs:fullTransport}
-\begin{multline} \label{eq:fullTransportVol}
-  \vec{\Omega}\vd \grad \psi(\vec{x}, \vec{\Omega})
-  + \sigma(\vec{x}) \psi (\vec{x}, \vec{\Omega})
-\\ =
-  \frac{\sigma_s(\vec{x})}{4\pi} \int_{4\pi} \psi(\vec{x},\vec{\Omega}')
-  \ud\Omega' + \frac{q(\vec{x})}{4\pi}
-  \equiv \frac{1}{4\pi} Q(\vec{x}) \,,
-\end{multline}
-inside $\vec{x} \in V$, $\vec{\Omega} \in 4\pi$, with an incident boundary
-condition
-\begin{equation} \label{eq:fullTransportBndy}
-  \psi(\vec{x}, \vec{\Omega}) = \psi^b(\vec{x}, \vec{\Omega}) \,,
- \quad \vec{x} \in \partial V, \ \vec{\Omega} \vd \vec{n} < 0\,.
-\end{equation}
-\end{subequations}
 
@@
 
-Talk about Dayman paper and how I'm extending that, to include neural nets
-first and to eval mucking up the training data second.
+This work begins by replicating the training and test sets used in ref (cite
+Dayman). The study does evaluate the imapact of randomly introduced error of
+varying amounts on the ability of the algorithms to correctly predict the
+burnup and reactor type. It also varies the amount of nuclides given as
+training to the algorithms (a full list, a short list determined by PCA, and
+fission products only). However, machine learning algorithms are heavily
+dependent on the inputs and parameters given to them, such as training set
+sizes, learning rates, etc. Additionally, it does not consider more advanced
+aglorithms. This paper adds in the comparison of the simpler algorithms to
+neural nets. It also presents learning curves for each algorithm to investigate
+the effect of the training set size on prediction error. For the classification
+of reactor type, ROC curves are also presented to evaluate prediction and
+generalization strength. For the regression of burnup, ??? pairwise t test?.
+Thus, I will first investigate these algorithms with no error introduced, but
+will extend the validation from a predetermined test set to be randomly
+partitioned cross validation.  And the next study will introduce error by
+limiting the nuclides to only those that can be measured with a gamma
+spectrometer (future work).
 
@@
 
-Preprocessing, kNN, ridge, neural net - these are all discussed above, but can
-talk deets + parameters here
+NN classifier+regressor, ridge classification + regression, neural net
+classification + regression- these are all discussed above, but can talk deets
++ parameters here
 
@@
 
-Classification Training Error: Predetermined test set for training error,
-perhaps can make a learning curve (maybe not until there is a variable
-train/test set for cross validation)
+Using predetermined test set:
 
-Classification and Regression Error Confidence: Confidence intervals on error
-(helps understand true error versus just the sample error), Test set must be >
-30 instances, Can easily calculate N\% confidence interval.
+1. Learning curves: For a given (randomly chosen) training set size for each
+algorithm, run several trainings and predictions. This helps determine if we
+are over or under training and each algorithm's robustness to over/under
+fitting.
 
-Generalizability should be evaluated (learning more about this with cs760
-project)
+2. Classification Training Error: Precision/recall values / ROC curves for the
+3 classes or maybe comparing algorithms to each other for each class
+
+3. Regression Training Error: Confidence intervals on predictions to understand
+true error versus sample error Test set must be > 30 instances, Can easily
+calculate N\% confidence interval.
+
+Using cross-validation instead:
+
+1. redo learning curve study without having to run several trainings and
+predictions for each set (bc that's already done!)
+
+2. redo precision/recall study
+
+3. redo confidence interval study
 
@@
 
-Options for comparison of algorithms: Comparing classification of 2 classes on
-same ROC plot with multiple ML systems, Scatter plots, Pairwise t-tests
+Maybe can delete this if I already discuss it in the above section. Options for
+comparison of algorithms: Comparing classification of 2 classes on same ROC
+plot with multiple ML systems, Scatter plots, Pairwise t-tests.
 
@@
 \section{Results and Analysis}
-The results were interesting, so interesting in fact that we have decided to
-present them here.
+
+So far, should include, with all 4 algorithms shown in each plot:
+1. Learning curves  + CV curve
+2. ROC plots + CV plot
+3. Confidence interval tables + CV table
+
+For the above categories in the validation section, can show plots directly
+next to each other (for 1, 2, 3) to hopefully show that cross validation
+provides better generalizability.  Future work would be to investigate the
+number of cross validation folds.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection Goes Here}
-The user must manually capitalize initial letters of a subsection heading.
-
-For those who like equations in their papers, \LaTeX\ is a good choice. Here is
-an equation for the Marshak diffusion boundary condition:
-\begin{equation} \label{eq:marshak}
-  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
-\end{equation}
-If we so choose, we can effortlessly reference the equation later.
-
-Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
-vacuum boundary condition:
-\begin{equation*}
-  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
-\end{equation*}
-The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
-an extrapolation distance of about $0.71045$.
-
-Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
-document. Always place figures after they are referenced so as not to throw
-off the reader. You can use symbols and different line styles to help
-differentiate your results, especially if they are printed in black and white.
-Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
-solution, solid lines \verb|-| for the new method's solutions, and dotted lines
-\verb|:| for existing inaccurate methods.
-\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
-  \centering
-  \includegraphics{example_figure}
-  \caption{Captions are flush with the left.}
-  \label{fig:voltage}
-\end{figure}
-
-Later on, we can include a table, even one that spans two columns such as
-Table~\ref{tab:widetable}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table*}[htb]
-  \centering
-\begin{tabular}{llllllllll}\toprule
-      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
-      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
-      $\varepsilon$      & $N_\text{it}$
-\\ \midrule
-$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
-\\
-$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
-\\
-$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
-\\
-$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
-\\
-\bottomrule
-\end{tabular}
-  \caption{This is an example of a really wide table which might not normally
-  fit in the document.}
-  \label{tab:widetable}
-\end{table*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Notice how the table reference uses a Roman numeral
-for its numbering scheme, whereas the figure reference uses an Arabic numeral.
-For one-column tables, use the \verb|table| environment; two-column tables use
-\verb|table*|. The same applies to figures.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Another Subsection}
-Excessive sectioning in a three-page document is discouraged, but here are more
-subsections to demonstrate compliance with the ANS formatting guidelines.
-
-\subsubsection{Third-level Heading}
-This subsubsection shows compliance with the ANS-specified standard. This level
-of heading should be used rarely.
-
-\subsubsection{Another Such Heading}
-And, if you really think you need a third-level heading, you should make sure
-that your subsection needs at least two of them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-The included ANS style file and this clear example file are a panacea for
-the hours of headache that invariably results from formatting a document in
-Microsoft Word.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Appendix}
-
-Numbering in the appendix is different:
-\begin{equation} \label{eq:appendix}
-  2 + 2 = 5\,.
-\end{equation}
-and another equation:
-\begin{equation} \label{eq:appendix2}
-  a + b = c\,.
-\end{equation}
+%\subsection{Subsection Goes Here}
+%The user must manually capitalize initial letters of a subsection heading.
+%
+%For those who like equations in their papers, \LaTeX\ is a good choice. Here is
+%an equation for the Marshak diffusion boundary condition:
+%\begin{equation} \label{eq:marshak}
+%  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
+%\end{equation}
+%If we so choose, we can effortlessly reference the equation later.
+%
+%Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
+%vacuum boundary condition:
+%\begin{equation*}
+%  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
+%\end{equation*}
+%The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
+%an extrapolation distance of about $0.71045$.
+%
+%Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
+%document. Always place figures after they are referenced so as not to throw
+%off the reader. You can use symbols and different line styles to help
+%differentiate your results, especially if they are printed in black and white.
+%Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
+%solution, solid lines \verb|-| for the new method's solutions, and dotted lines
+%\verb|:| for existing inaccurate methods.
+%\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
+%  \centering
+%  \includegraphics{example_figure}
+%  \caption{Captions are flush with the left.}
+%  \label{fig:voltage}
+%\end{figure}
+%
+%Later on, we can include a table, even one that spans two columns such as
+%Table~\ref{tab:widetable}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{table*}[htb]
+%  \centering
+%\begin{tabular}{llllllllll}\toprule
+%      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
+%      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
+%      $\varepsilon$      & $N_\text{it}$
+%\\ \midrule
+%$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
+%\\
+%$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
+%\\
+%$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
+%\\
+%$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
+%\\
+%\bottomrule
+%\end{tabular}
+%  \caption{This is an example of a really wide table which might not normally
+%  fit in the document.}
+%  \label{tab:widetable}
+%\end{table*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Notice how the table reference uses a Roman numeral
+%for its numbering scheme, whereas the figure reference uses an Arabic numeral.
+%For one-column tables, use the \verb|table| environment; two-column tables use
+%\verb|table*|. The same applies to figures.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Another Subsection}
+%Excessive sectioning in a three-page document is discouraged, but here are more
+%subsections to demonstrate compliance with the ANS formatting guidelines.
+%
+%\subsubsection{Third-level Heading}
+%This subsubsection shows compliance with the ANS-specified standard. This level
+%of heading should be used rarely.
+%
+%\subsubsection{Another Such Heading}
+%And, if you really think you need a third-level heading, you should make sure
+%that your subsection needs at least two of them.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Conclusions}
+%
+%The included ANS style file and this clear example file are a panacea for
+%the hours of headache that invariably results from formatting a document in
+%Microsoft Word.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\appendix
+%\section{Appendix}
+%
+%Numbering in the appendix is different:
+%\begin{equation} \label{eq:appendix}
+%  2 + 2 = 5\,.
+%\end{equation}
+%and another equation:
+%\begin{equation} \label{eq:appendix2}
+%  a + b = c\,.
+%\end{equation}
 

### 2017-05-18 paper.tex
updates

@@
 
-Workflow for determining SNF quantities of interest: measure material, use
-isotope content and/or isotope ratios to determine things like reactor type,
-fuel type and enrichment at beginning of irradiation, cooling time, burnup.
-
-After the material characteristics are measured, they are matched in a
-forensics database that includes some or all this information for
+This study focuses on non-detonated materials, specifically, spent nuclear
+fuel. It is important to determine if some intercepted material is from a
+commercial fuel cycle or if it is meant for weapons production (and where the
+material was obtained from).  Workflow for determining SNF quantities of
+interest: measure material, use isotope content and/or isotope ratios to
+determine things like reactor type, fuel type and enrichment at beginning of
+irradiation, cooling time, burnup.  (Classification, Characterization,
+Interpretation (Analysis), Reconstruction (Attribution) - from the New Nuclear
+Forensics book) (Char methods to get isotopic ratios or use S/ML, Interp
+examples) After the material characteristics are measured, they are matched in
+a forensics database that includes some or all this information for
 pre-existing/pre-measured SNF. These databases are kept by individual
@@
 
-(Classification, Characterization, Interpretation (Analysis), Reconstruction
-(Attribution) - from the New Nuclear Forensics book) (Char methods to get
-isotopic ratios or use S/ML, Interp examples)
-
-To accomplish this, (we want to see if) it is possible to limit material
+So a lofty goal for the forensics community would be to develop methods that
+provide instantaneous information that is reliable enough to guide an
+investigation (e.g., within 24 hours). Fast, accurate measurements to provide
+isotopic ratios to calculate the above-mentioned fuel parameters of interest
+would provide this. To accomplish/mimic this, we wil limit material
 measurements to rapid-result nondestructive analysis, such as gamma ray
@@
 Additionally, within gamma spec techniques (field vs lab), uncertainties can
-also vary.
-
-(Something on goal to get as much info as possible within 24 h of interception
-or device detonation).
+also vary. However, using a well-trained machine-learned model may be able to
+overcome these inherent issues with gamma spec. The current and future work of
+this study is designed with this in mind.
 
@@
 and machine learning tools have been used to both classify spent fuel (reactor
-type, fuel type) and predict values such as burnup, initial enrichment, or
-cooling time (regression).
+type, fuel type) and provide regression values such as burnup, initial
+enrichment, or cooling time.
 
-Add some real ML background here
+In order to validate my own work, I am choosing some machine learning
+approaches based on previous work (cite) while also extending to more complex
+algorithms. Thus, this paper investigates four algorithms: nearest neighbors
+using two different distance metrics, ridge regression, and a neural network.
+These algorithms can all be modified to do either classification or regression.
+
+In this paper, run through em all quickly, but in prelim make each into it's
+own section.
 
@@
 This work begins by replicating the training and test sets used in ref (cite
-Dayman). The study does evaluate the imapact of randomly introduced error of
+Dayman). Replicate table of training instances perhaps. Discuss how this covers
+both short and long fuel cycles and their relation to weapons versus commercial
+use.
+
+The study does evaluate the imapact of randomly introduced error of
 varying amounts on the ability of the algorithms to correctly predict the

### 2017-07-26 paper.tex
progress from last 2 days

@@
 \usepackage{microtype} % improves typography for PDF
+\usepackage{todonotes}
 
@@
 material or the detonation of a dirty bomb, it is necessary to learn as much as
-possible about the source of the materials in a timely manner. To characterize
-the materials, both radiological methods (e.g., gamma spectroscopy) and
-ionization methods (e.g., mass spectroscopy) are used to determine isotopic
-ratios, chemical compounds, or trace elements. Although each category has a
-multitude of techniques within it, the main tradeoff is between time/cost and
-amount of information gained.
+possible about the source of the materials in a timely manner. In the case of
+non-detonated special nuclear material, knowing the reactor parameters that
+produced it can point investigators in the right direction in order to
+determine the chain of custody of the interdicted material. Determining these
+parameters (e.g., cooling time, burnup) requires first characterizing and
+calculating certain isotopic ratios, chemical compounds, or trace elements.
+Both radiological methods (e.g., gamma spectroscopy) and ionization methods
+(e.g., mass spectroscopy) measure these quantities. Although both measurement
+techniques have a multitude of techniques within them and thus varying
+strengths and weaknesses, the main tradeoff is between time/cost and amount of
+information gained.
 
 The results of these analytic techniques are then compared against existing
-databases to determine the origin of the nuclear material(s). These databases
-are highly multidimensional, and furthermore, are rife with missing data
-entries and inconsistent uncertainties. Direct comparison between measurement
-results and a database therefore may not yield accurate results. Fortunately,
-machine learning algorithms can be explored to create a model from a database
-to "fill between the lines".  Additionally, having a machine-learned model may
+databases to obtain the desired reactor parameters. These databases are highly
+multidimensional, and furthermore, are rife with missing data entries and
+inconsistent uncertainties. Direct comparison between measurement results and a
+database therefore may not yield accurate results. Thus, computational
+techniques have been developed by nuclear engineers to calculate the parameters
+relevant to nuclear forensics analysis. \todo{Cite inverse stuff, possibly
+adjoint?} Another approach requiring minimal domain knowledge is the use of
+statistical methods via machine learning algorithms. These algorithms can
+create a model using the database entries that enables "filling between the
+lines" of its entries. Additionally, having a machine-learned model may
 overcome the above challenges of multidimensionality, missing data, and
@@
 extremely fast results cheaply, it only measures certain radiological signals
-and is influenced by many environmental (or storage) factors. In the simulation
-and machine learning paradigm, we need to determine what exactly is needed to
-train a machine-learned model. Can the algorithm overcome the deficiencies of
-gamma detection and still provide useful results? Or does it need more
-information, e.g., exact isotopics?
+and is influenced by many environmental factors, storage, and self-attenuation.
 
-Thus, ultimately, the goal is to answer the question \textit{How does the
-ability to determine forensic-relevant spent nuclear fuel attributes degrade as
-less information is available?}. But first, we must establish some baseline
-expectations and algorithms to use. This work is based off previous work on the
-subject (cite Dayman), and expands upon it by also evaluating a more advanced
-machine learning algorithm: neural nets. Below is a more in depth discussion of
-nuclear forensics and how machine learning can contribute to this research
-area. After that, an experimental design is outlined. Lastly, the results are
-presented and discussed.
+\todo{assumes previous section on fuel cycle sim} In the simulation and machine
+learning paradigm, we need to determine what exactly is needed to train a
+machine-learned model. Of interest to an entity trying to create a weapon is
+partially irradiated fuel if they have plutonium separations capabilities or
+any radioactive substance in the case of a dirty bomb. Addressing the former,
+we used a set of simulations of spent nuclear fuel at different burnups and
+cooling times.
+
+Can the algorithm overcome the deficiencies of gamma detection and still
+provide useful results? Or does it need more information, e.g., exact
+isotopics? Thus, ultimately, the goal is to answer the question \textit{How
+does the ability to determine forensic-relevant spent nuclear fuel attributes
+degrade as less information is available?}.
+
+But first, we must establish some baseline expectations of reactor parameter
+prediction and algorithms to use.  This work is based off previous work on the
+subject (cite Dayman) regarding machine learning performace with respect to
+information reduction, and expands upon it by also evaluating a more advanced
+machine learning algorithm: support vector regression. Below is a more in depth
+discussion of nuclear forensics and how machine learning can contribute to this
+research area. After that, an experimental design is outlined. Lastly, the
+results are presented and discussed.
 
@@
 
-Nuclear forensics is the analysis and interpretation to determine the history
-of nuclear material, whether that be intercepted spent nuclear fuel, uranium
-ore concentrate, or the debris from an exploded nuclear device. Technical
-nuclear forensics focuses on the characterization and interpretation of those
-results. This, in combination with intelligence data, aids in material
-attribution, which is the overall goal of nuclear forensics.
+The process of nuclear forensics includes the analysis and interpretation of
+nuclear material to determine its history, whether that be intercepted spent
+nuclear fuel, uranium ore concentrate, or the debris from an exploded nuclear
+device. After the technical portion is complete, intelligence data can be used
+to aid in material attribution; this is the overall goal of nuclear forensics.
 
@@
 commercial fuel cycle or if it is meant for weapons production (and where the
-material was obtained from).  Workflow for determining SNF quantities of
-interest: measure material, use isotope content and/or isotope ratios to
-determine things like reactor type, fuel type and enrichment at beginning of
-irradiation, cooling time, burnup.  (Classification, Characterization,
-Interpretation (Analysis), Reconstruction (Attribution) - from the New Nuclear
-Forensics book) (Char methods to get isotopic ratios or use S/ML, Interp
-examples) After the material characteristics are measured, they are matched in
-a forensics database that includes some or all this information for
-pre-existing/pre-measured SNF. These databases are kept by individual
-countries, and a given database will have widely varying uncertainty depending
-where the material was measured as well as missing data in some fields.
-Therefore, matching can be difficult.
+material was obtained from).
 
-So a lofty goal for the forensics community would be to develop methods that
+\todo{Workflow for determining SNF quantities of interest}: measure material, use
+isotope content and/or isotope ratios to determine things like reactor type,
+fuel type and enrichment at beginning of irradiation, cooling time, burnup.
+(Classification, Characterization, Interpretation (Analysis), Reconstruction
+(Attribution) - from the New Nuclear Forensics book) (Char methods to get
+isotopic ratios or use S/ML, Interp examples) After the material
+characteristics are measured, they are matched in a forensics database that
+includes some or all this information for pre-existing/pre-measured SNF. These
+databases are kept by individual countries, and a given database will have
+widely varying uncertainty depending where the material was measured as well as
+missing data in some fields.  Therefore, matching can be difficult.
+
+A lofty goal for the forensics community would be to develop methods that
 provide instantaneous information that is reliable enough to guide an
-investigation (e.g., within 24 hours). Fast, accurate measurements to provide
-isotopic ratios to calculate the above-mentioned fuel parameters of interest
-would provide this. To accomplish/mimic this, we wil limit material
-measurements to rapid-result nondestructive analysis, such as gamma ray
-spectroscopy. Gamma spec gives less information at a higher uncertainty than
-the near perfect results of some destructive mass spec techniques, like TIMS.
-Additionally, within gamma spec techniques (field vs lab), uncertainties can
-also vary. However, using a well-trained machine-learned model may be able to
-overcome these inherent issues with gamma spec. The current and future work of
-this study is designed with this in mind.
+investigation (e.g., within 24 hours). Fast measurements to provide isotopic
+ratios to calculate the above-mentioned fuel parameters of interest would
+provide this via some form of a handheld detector that measures gamma spectra.
+However, while this nondestructive analysis is rapid, it is also difficult to
+evaluate because of the presence of overlapping peaks.  Thus, gamma spectra
+give less information at a higher uncertainty than the near-perfect results of
+some destructive mass spectroscopy techniques, like TIMS\todo{find MS technique
+and citation}.  Additionally, within gamma spectroscopy techniques (e.g., field
+vs. lab detectors), uncertainties can vary significantly because of the
+detector reponse, environment, storage, electronics, etc.  However, using a
+well-trained machine-learned model may be able to overcome these inherent
+issues with gamma spectra. The current and future work of this study is
+designed with this in mind.
 
@@
 Given imperfect data with varying amounts of uncertainty as well as the
-required comparison to imperfect databases, many have begun considering
-artificial intelligence approaches to nuclear forensics problems, such as
-implementing searching algorithms for database comparison and machine learning
-for determining spent fuel characteristics (cite all). A variety of statistical
-and machine learning tools have been used to both classify spent fuel (reactor
-type, fuel type) and provide regression values such as burnup, initial
-enrichment, or cooling time.
+required comparison to highly multidimensional databases with missing entries,
+many have begun considering computational approaches to nuclear forensics
+problems, such as the INDEPTH\todo{cite and discuss}
 
-In order to validate my own work, I am choosing some machine learning
-approaches based on previous work (cite) while also extending to more complex
-algorithms. Thus, this paper investigates four algorithms: nearest neighbors
-using two different distance metrics, ridge regression, and a neural network.
-These algorithms can all be modified to do either classification or regression.
+Another approach utilizes artificial intelligence to solve nuclear forensics
+problems, such as implementing searching algorithms for database comparison and
+machine learning for determining spent fuel characteristics \todo{cite
+hungarian guy, viz guy, dayman paper}.  A variety of statistical and machine
+learning tools have been used to characterize spent fuel by predicting
+categories or labels (reactor type, fuel type) as well as predicting values
+(burnup, initial enrichment, or cooling time) The former uses classification
+algorithms and the latter uses regression algorithms. Many algorithms can be
+applied to both cases.
 
-In this paper, run through em all quickly, but in prelim make each into it's
-own section.
+A typical (supervised) machine learning workflow would take a set of training
+data with labels or values inserted into some statistical learner, calculate
+some objective, minimize or maximize that objective, and provide some model
+based on that output. Then a test set (with known values) is provided to the
+model so that its performance can be evaluated and finalized. After model
+finalization, a user can provide a single instance and a value can be predicted
+from that. \todo{insert ML schematic}
+
+To obtain reliable models, one must 1. choose/create a training set carefully
+and 2. study the impact of various algorithm parameters on the error. Many
+algorithms are developed on an assumption that the training set will be
+independent and identially distributed (i.i.d.). [Aside: there are ways to
+handle skewed data sets] This is important so that the model does not overvalue
+or overfit a certain area in the training space. Additionally, algorithm
+performance (or error) can be optimized with respect to training set size,
+number of features, or algorithm parameters (regularization terms, etc).  These
+are known as diagnostic plots. When plotting the training and testing error
+with respect to the number of instances, this is known as a learning curve.
+When plotting these errors with respect to the number or features or algorithm
+parameters, this is known as a validation curve. \todo{insert example
+diagnostic plot?}
+
+Algorithm choice is usually based on what is being predicted and intuition
+regarding strengths and weaknesses.  For the sake of comparison (i.e. weak
+validation), some machine learning approaches here are based on previous
+work\todo{cite dayaman} while also extending to a more complex model via an
+algorithm that is known to handle highly dimensional data sets well. Thus, this
+paper investigates three regression algorithms: nearest neighbor, ridge, and
+support vectors.
+
+\subsubsection{Nearest Neighbor Regression}
+
+\todo{intro, alg math, etc}
+
+Nearest neighbor regression calculates a value based on the instance that is
+closest to it. The metrics for distance differ, but in this study, Euclidian
+distance was used. There is no learning in this regression, per se; the
+training set populates a space and the testing set is compared directly to
+that.
+
+\subsubsection{Ridge Regression}
+
+\todo{intro, alg math, etc}
+
+Ridge regression utilizes normal linear least squares regression, but with a
+parameter that penalizes correct answers to prevent overfitting, which is
+referred to as regularization.
+
+\subsubsection{Support Vector Regression}
+
+\todo{add schematics from pres...or diff pics with the math?}
+
+Support vector regression (SVR) is an extension of the popular classification
+algorithm, support vector machine (SVM).  This algorithm was chosen because of
+its ability to handle highly dimensional data well, which in this study is
+approximately 300 features.
+
+SVM classifies two classes by determining an optimal hyperplane, given by wx+b,
+between them.  As seen in Figure ?, the algorithm evaluates the quality of the
+line that separates two classes by maximizing the width of the margin given the
+contraints surrounding the line.  Some problems are not linearly separable, and
+thus a penalty term is introduced to allow for misclassifications. As shown in
+Figure ?, the algorithm then simultaneously minimizes the misclassifications
+while maximizing the margin.
+
+This can be extended easily to multidimensional analysis via what is called the
+\textit{kernel trick}.  First, using a nonlinear kernel function maps the data
+into higher dimensional feature space. Then the algorithm can find a linear
+separation in this space, as shown in Figure ?. Further, this can be upgraded
+from classification to SVR by doing similar math but instead minimizing the
+margin, as shown in Figure ?.
+
+The kernel chosen for this study is the Gaussian radial basis function, shown
+below. This has two tuneable parameters, gamma and C. Gamma influences the
+width of influence of individual training instances, and strongly affects the
+fitting of the model. Low values correspond to underfitting because the
+instances have too large of a radius (low influence) and high values correspond
+to overfitting because the instances have a small radius (high influence).
+
+The C parameter also affects the fitting of the model by allowing more or less
+support vectors, corresponding to more or less misclassification. A lower C
+smooths the surface of the model by allowing more misclassifications, whereas a
+higher C classifies more training examples by allowing fewer
+misclassifications. Too low and too high of a C can cause under- and
+overfitting, respoectively.
+
+Since there is a tradeoff of fitting strength provided by both parameters, it
+is common to run the algorithm on a logarithmic grid from 10'-3 to 10'3 for
+each parameter. If plotted on a heatmap of accuracies given gamma and C, there
+will be a diagonal of ideal combinations that emerges. The lowest of these is
+usually chosen.
 
@@
 
-This work begins by replicating the training and test sets used in ref (cite
-Dayman). Replicate table of training instances perhaps. Discuss how this covers
-both short and long fuel cycles and their relation to weapons versus commercial
-use.
+\subsection{Testing and Training Sets}
 
-The study does evaluate the imapact of randomly introduced error of
-varying amounts on the ability of the algorithms to correctly predict the
-burnup and reactor type. It also varies the amount of nuclides given as
-training to the algorithms (a full list, a short list determined by PCA, and
-fission products only). However, machine learning algorithms are heavily
-dependent on the inputs and parameters given to them, such as training set
-sizes, learning rates, etc. Additionally, it does not consider more advanced
-aglorithms. This paper adds in the comparison of the simpler algorithms to
-neural nets. It also presents learning curves for each algorithm to investigate
-the effect of the training set size on prediction error. For the classification
-of reactor type, ROC curves are also presented to evaluate prediction and
-generalization strength. For the regression of burnup, ??? pairwise t test?.
-Thus, I will first investigate these algorithms with no error introduced, but
-will extend the validation from a predetermined test set to be randomly
-partitioned cross validation.  And the next study will introduce error by
-limiting the nuclides to only those that can be measured with a gamma
-spectrometer (future work).
+This work begins by simulating the training and test sets described in ref
+(cite Dayman). As with the previous work, this will be done using SCALE 6.2
+\todo{cite}. Specifically, the ARP module of the activation and depletion code
+ORIGEN was used. \todo{add deets and citations}
 
-\subsection{Algorithms Used}
+\todo{show table of training set space}
+The parameters of the training set are defined as follows. A smaller burnup
+than is typical for spent fuel from a commercial reactor is used in the
+previous work likely because stolen fuel pins for weapons use would not likely
+be at the end of their lifetime, as the plutonium of interest has decreased by
+then. A truly i.i.d. training set would go beyond this, but this is purely for
+demonstration with a single use case in mind.
 
-NN classifier+regressor, ridge classification + regression, neural net
-classification + regression- these are all discussed above, but can talk deets
-+ parameters here
+\todo{show table of test set space}
+The previous work also used an external test set, designed to have values in
+between the trained values of burnup. This is implemented in this study but it
+is expected that cross-validation will better indicate the model performance.
+More specifically, using k-fold cross-validation is a common method to use in
+the application of machine learning to create more confidence in the resulting
+model. \todo{either explain or cite cross validation next}
 
-\subsection{Validation of Each Algorithm}
+\subsection{Information Reduction}
+
+The study does evaluate the impact of randomly introduced error of varying
+amounts on the ability of the algorithms to correctly predict the burnup. It
+also varies the amount of nuclides given as training to the algorithms (a full
+list, a short list determined by PCA, and fission products only). However,
+machine learning algorithms are heavily dependent on the inputs and parameters
+given to them, such as training set sizes, learning rates, etc. Additionally,
+it does not consider more advanced aglorithms. This paper adds in the
+comparison of the simpler algorithms to neural nets. It also presents learning
+curves for each algorithm to investigate the effect of the training set size on
+prediction error. For the classification of reactor type, ROC curves are also
+presented to evaluate prediction and generalization strength. For the
+regression of burnup, ??? pairwise t test?.  Thus, I will first investigate
+these algorithms with no error introduced, but will extend the validation from
+a predetermined test set to be randomly partitioned cross validation.  And the
+next study will introduce error by limiting the nuclides to only those that can
+be measured with a gamma spectrometer (future work).
+
+\subsection{Algorithm Evaluation}
 
@@
 
-2. Classification Training Error: Precision/recall values / ROC curves for the
-3 classes or maybe comparing algorithms to each other for each class
-
-3. Regression Training Error: Confidence intervals on predictions to understand
+2. Regression Training Error: Confidence intervals on predictions to understand
 true error versus sample error Test set must be > 30 instances, Can easily
@@
 
-2. redo precision/recall study
-
-3. redo confidence interval study
+2. redo confidence interval study
 
@@
 
-So far, should include, with all 4 algorithms shown in each plot:
-1. Learning curves  + CV curve
-2. ROC plots + CV plot
-3. Confidence interval tables + CV table
+1. Learning curve
+2. Validation curve
+3. Random error curve
 
-For the above categories in the validation section, can show plots directly
-next to each other (for 1, 2, 3) to hopefully show that cross validation
-provides better generalizability.  Future work would be to investigate the
-number of cross validation folds.
+Want to show the difference between test/train error plots and CV/train error
+plots.  Determine some argument that perfers the latter. For the above
+categories in the validation section, can show plots directly next to each
+other (for 1, 2, 3) to hopefully show that cross validation provides better
+generalizability.
+
+\todo{Discuss understanding confidence intervals in predictions.}
+
+\section{Notes from Meeting Regarding Signatures}
+\subsection{U Processing Signatures}
+Isotopics provide the following: 
+1. Enrichment indicators
+2. radiochronometers provide info on separations, metal, etc
+3. heterogeneity of isotopics provide info on blending
+4. stable isotopes/atmosphere 
+Chemical characteristics ask: 
+1. chemical products and trace contaminants @ each step
+2. aging in different atmospheres
+Processing Signatures in particular:
+Shifting focus towards U metals
+
+\subsection{Pu Processing Signatures}
 

### 2017-07-26 paper.tex
wed afternoon updates

@@
 
-\todo{intro, alg math, etc}
+\todo{intro, alg math, explain distance metrics}
 
@@
 
-\todo{intro, alg math, etc}
+\todo{intro, alg math, explain regularization}
 
@@
 
-The study does evaluate the impact of randomly introduced error of varying
-amounts on the ability of the algorithms to correctly predict the burnup. It
-also varies the amount of nuclides given as training to the algorithms (a full
-list, a short list determined by PCA, and fission products only). However,
-machine learning algorithms are heavily dependent on the inputs and parameters
-given to them, such as training set sizes, learning rates, etc. Additionally,
-it does not consider more advanced aglorithms. This paper adds in the
-comparison of the simpler algorithms to neural nets. It also presents learning
-curves for each algorithm to investigate the effect of the training set size on
-prediction error. For the classification of reactor type, ROC curves are also
-presented to evaluate prediction and generalization strength. For the
-regression of burnup, ??? pairwise t test?.  Thus, I will first investigate
-these algorithms with no error introduced, but will extend the validation from
-a predetermined test set to be randomly partitioned cross validation.  And the
-next study will introduce error by limiting the nuclides to only those that can
-be measured with a gamma spectrometer (future work).
+This study evaluates the impact of randomly introduced error of varying amounts
+on the ability of the algorithms to correctly predict the burnup. Thus, first
+investigated are the three algorithms with no error introduced, and next with
+the random errors applied uniformly to each nuclide vector. Since error in a
+nuclide vector is not random, in fact it is systematic and dependent on a
+number of known sources of uncertainty, the next study will introduce error by
+limiting the nuclides to only those that can be measured with a gamma
+spectrometer (future work).
 
@@
 
-Using predetermined test set:
+Additionally, machine learning algorithms are heavily dependent on the inputs
+and parameters given to them, such as training set sizes, learning rates,
+regularization, etc. To evaluate the performance or tweak the model from an
+algorithm, diagnostic plots will be used. Learning and validation curves will
+indicate how the models are performing, initially both with respect to the
+testing error and the cross validation error. As previously mentioned, these
+two errors are to be compared to the training error to understand the
+prediction and generalization strength with respect to training set size and
+the algorithm parameters governing model complexity.
 
-1. Learning curves: For a given (randomly chosen) training set size for each
-algorithm, run several trainings and predictions. This helps determine if we
-are over or under training and each algorithm's robustness to over/under
-fitting.
+The learning curves were obtained as follows. For a given (randomly chosen)
+training set size between 15 and 100\% of the total data set, several training
+and prediction rounds were performed. The repetition for obtaining the testing
+error is the same value as the \textit{k} in k-fold cross validation. The
+testing error scenario averages the values of the obtained errors whereas the
+k-fold cross-validation performs this automatically.  The validation curves
+were obtained as follows. For a given parameter, the value of the parameter is
+varied and \textit{k} training and prediction phases are completed, and their
+errors averaged. Again, for k-fold cross-validation, these errors are already
+averaged. The learning curves help determine if we are over- or under-training.
+The validation curves help determine the optimal way to be robust to over- and
+under-fitting.
 
-2. Regression Training Error: Confidence intervals on predictions to understand
-true error versus sample error Test set must be > 30 instances, Can easily
-calculate N\% confidence interval.
-
-Using cross-validation instead:
-
-1. redo learning curve study without having to run several trainings and
-predictions for each set (bc that's already done!)
-
-2. redo confidence interval study
+% confidence interval stuff for future paper
+%2. Regression Training Error: Confidence intervals on predictions to understand
+%true error versus sample error Test set must be > 30 instances, Can easily
+%calculate N\% confidence interval.
+%2. redo confidence interval study
 
